---
title: "EDA"
output:
  html_document:
    df_print: paged
---

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(naniar)
library(cluster)
library(factoextra)
library(dygraphs)
library(xts)
```

```{r}
owid_covid_data <- read_csv("~/Documents/Programming/Citadel-Datathon/Datathon Materials/1_owid/owid-covid-data.csv",
                            col_types = paste(c(rep("c",3),"D",rep("d",29),"c",rep("d",18)),collapse = ""))
```


# Clustering

Let's try to find similar countries based on time invariant variables.

```{r}
# so these "country metadata columns" are constant where available
test <- owid_covid_data%>%
  group_by(location)%>%
  summarise(across(37:51,var, na.rm=TRUE))
sum(test[,-1], na.rm = TRUE) # 0

# clustering
test <- owid_covid_data%>%
  select(location, 38:52)%>%
  group_by(location)%>%
  slice(1)

sum(is.na(test)) # 361
vis_miss(test)

```


```{r}
which.max(test%>%is.na()%>%rowSums()) # 80 = "international"
```

Let's drop "International" as location.

```{r}
# remove "International"
test <- owid_covid_data%>%
  select(location, 38:52)%>%
  group_by(location)%>%
  slice(1)%>%
  filter(location != "International")%>%
  column_to_rownames("location")

# Gower can handle missing values
d <- daisy(test,metric = "gower")
fit <- hclust(d)
plot(fit, labels = test$location, cex=0.5)
rect.hclust(fit,k=5,border = "red")
```

```{r}
# this one suggests there might be 3 clusters?
fviz_dist(d, gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07"))+
  theme(axis.text = element_text(size = 3))
```

The "red cross" in the bottom left corner is Monaco, which was also a cluster on its own in the previous plot. Also, Monaco is probably not too representative. Let's drop it and repeat the clustering.

```{r}
test <- test%>%rownames_to_column()%>%
  filter(rowname != "Monaco")%>%
  column_to_rownames("rowname")

# Now cluster again
d <- daisy(test,metric = "gower")
fit <- hclust(d)

plot(fit, cex=0.3)
rect.hclust(fit,k=5,border = "red")
```

Let's grab those 5 groups.

```{r}
test$group <- cutree(fit,k=5)
test <- test%>%rownames_to_column("location")
```


# Time series

Group 4 are generally highly developed countries, althoug it is to be noted that this group does not include the US, Canada not Australia.

```{r}
test%>%
  filter(group==4)%>%
  select(location)
```



```{r}
df <- 
  owid_covid_data%>%
  left_join(test%>%select(location, group), by = "location")%>%
  filter(group == 4)%>%
  filter(population > 5e6)

df%>%
  ggplot(aes(x = date, y = new_cases_smoothed_per_million, color = location))+
  geom_line()+
  labs(title = "New Cases per Million smoothed")
```

```{r}
df%>%
  ggplot(aes(x = date, y = new_deaths_smoothed_per_million, color = location))+
  geom_line()+
  labs(title = "New Deaths per Million smoothed")
```

```{r}
df%>%
  ggplot(aes(x = date, y = stringency_index, color = location))+
  geom_line()+
  labs(title = "Stringency")
```



```{r, fig.width=8}
country <- "Sweden"

xts(x = df%>%filter(location == country)%>%select(stringency_index, new_cases_smoothed_per_million),
    order.by = df%>%filter(location == country)%>%pull(date))%>%
  dygraph()%>%
  dySeries("stringency_index", axis = 'y2')%>%
  dyAxis("y", label = "New cases per million (smoothed)")%>%
  dyAxis("y2", label = "Stringency Index", valueRange = c(0,100))%>%
  dyRangeSelector()%>%
  dyCrosshair("both")
```

